sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-12-02 doc/paper.Rmd
add notes

@@

sdmTMB fits GLMMs to spatial or spatiotemporal geostatistical data.
Geostatistical data simply refers to data observed at specific spatial coordinates. The data can be collected across discrete points in time. Areal data (data on polygons or grids) may be analyzed using other spatial models, including conditional autoregressive models (CAR). sdmTMB can also fit models with areal data if each polygon has an associated centroid (observations are treated as replicates). A benefit of the sdmTMB approach over CAR or similar models is the parameters describing spatial covariance can be estimated (e.g. range).

@@

@@

<!-- Future directions para? -->
<!-- multivariate, but only if can keep interface simple -->
<!-- optimized Bayes sampling -->
<!-- residual diagnostics -->
<!-- zero inflation -->
<!-- MVT? time-varying epsilon -->

@@

>>>>

<!-- TODO cite: -->
<!-- Dormann, C.F., McPherson, J.M., Araújo, M.B., Bivand, R., Bolliger, J., Carl, G., Davies, R.G., Hirzel, A., Jetz, W., Kissling, W.D., Kühn, I., Ohlemüller, R., Peres‐Neto, P.R., Reineking, B., Schröder, B., Schurr, F.M., and Wilson, R. (n.d.). Methods to account for spatial autocorrelation in the analysis of species distributional data: a review. Ecography 30(5): 609–628. doi:10.1111/j.2007.0906-7590.05171.x. -->

@@

<!-- or areal/grid data with centroids instead of CAR/SAR models -->
<!-- add CAR/SAR to table -->
sdmTMB fits GLMMs to spatial or spatiotemporal geostatistical data.
Geostatistical data simply refers to data observed at specific spatial coordinates. The data can be collected across discrete points in time. Areal data (data on polygons or grids) may be analyzed using other spatial models, including conditional autoregressive models (CAR). sdmTMB can also fit models with areal data if each polygon has an associated centroid (observations are treated as replicates). A benefit of the sdmTMB approach over CAR or similar models is the parameters describing spatial covariance can be estimated (e.g. range).

@@

<!-- highlight where each one excels ... -->
<!-- no CV built into VAST or priors -->
<!-- discussion integrates across speed, familiarity and table 1 -->
<!-- results paragraph on Table 1 -->
<!-- familiarity is an important word here -->
<!-- add section on functionality comparison -->

@@

<!-- Future directions para? -->
<!-- multivariate, but only if can keep interface simple -->
<!-- optimized Bayes sampling -->
<!-- residual diagnostics -->
<!-- zero inflation -->
<!-- MVT? time-varying epsilon, non-gaussian fields -->

<!-- limitations -->

<!-- 3d models, continuous time, -->
<!-- areal models, neighbours or not, must be CAR or SAR, ... but if have centroid, could use sdmTMB... why not CAR/SAR, spatial process... not 1:1 relationship between distance and spatial correlation? neibour based only... -->
<!-- https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.701.1636&rep=rep1&type=pdf -->

@@

<!-- CAR/SAR models? and why not if have...-->
<!-- EOF paper from Jim-->
<!-- testing that has been done... -->
<!-- https://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.15149 -->
<!-- https://onlinelibrary.wiley.com/doi/abs/10.1111/gcb.15149 -->
